---
title: "COVID-19 Effects on the Real Estate Market"
author: "Joshua Kim, Daniel Choe, Ethan Gilman"
date: "05/18/2021"
output:
  rmdformats::readthedown:
    thumbnails: false
    highlight: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(tidyverse)
library(readxl)
library(robotstxt)
library(knitr)
library(janitor)
library(rvest)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(leaflet)
library(leaflet.extras)
library(maps)
library(lubridate)
```

```{r, echo=FALSE, include=FALSE}
#Importing Wrangled CSV Files for visualizations
BPvCM <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/BPvCM.csv")
CPIHPI <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/CPIHPI.csv")
SPHomePrice <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/SPHomePrice.csv")
BoxPlotData <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/BoxPlotData.csv")
FullTableLong <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/FullTableLong.csv")
linegraph_data <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/linegraph_data.csv")
scatter_data <- read_csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/scatter_data.csv")
```

# Introduction
The median homeowner has 40 times the household wealth of a renter, according to the 2019 Survey of Consumer Finances, a triennial survey that collects detailed accounts of households’ finances (Kushi 2020). Other studies show that homeownership leads to greater wealth-building in comparison to renting. Thus, real estate and homeownership is still one of the best ways to build wealth, but is it really a good investment and does it perform well throughout economic downturns? Specifically how did it perform throughout the COVID-19 pandemic?

Despite flashbacks of the 2008 financial crisis and recessionary concerns due to the pandemic, the real estate frenzy has reignited after a short downturn. Many houses are selling well over asking price due to heavy competition and single family home inventory is at ⅔ of last year’s based on Aug 2019 and 2020. Demand is growing while supply is decreasing, snowballing the real estate market into a hot, competitive market. This is increasing the barrier to entry as the amount of capital needed to purchase real estate is also growing.

We look to explore the single family real estate market and the impact of COVID-19 on the market. We specifically are looking at single family homes (SFH) because of the higher liquidity within the market, lowest barrier to entry, and because commercial data is not publicly available.

In order to examine such phenomena mid-pandemic, we seek to answer, _‘How did the COVID-19 pandemic impact the real estate industry?’ ‘How well has the real estate market performed compared to other markets?’ and ‘What explains this real estate market behavior?’_

# Overview of Analysis
In order to analyze the impact of Covid on the real estate market, we first turn to the impact on national sales volume regarding SFH. Then we seek any differences between average home prices of the US and certain states. Then we seek to understand the real estate on a more granular level by analyzing how prices were impacted by state. This will help us understand how overall the supply and demand of the market affected the prices and quantities that homes are selling for.

Once we realize the overall market effects, we examine how the real estate market has performed vs other markets. We do this by comparing the Home Price Index with the Consumer Price Index and the S&P 500. This will allow us to understand how the real estate market performed in comparison with the average cost of living and compared to the top 500 stocks in the US.

Then we try to understand what developments caused the market to react in the way that it did. Two common allegations are that it is because interest rates are low and because construction costs are high. Thus we plotted the interest rate and home price to understand the relationship between interest rates and home prices. Then, we compared the percentage change between construction material costs and building permits approved to examine the relationship between construction costs and newly built homes. 

# Data
Our data was extracted from a variety of sources such as the Federal Reserve Bank of St. Louis, US Census, and Zillow. 

The data for national home sales count and median home sale price came from Zillow.

[Zillow](https://www.zillow.com/research/data/)

The data for CPI, HPI, and S&P 500 came from the Federal Reserve Bank of St. Louis, also known as the FRED.

[HPI](https://fred.stlouisfed.org/series/CSUSHPINSA)
[CPI](https://fred.stlouisfed.org/series/USACPIALLMINMEI)
[S&P 500](https://fred.stlouisfed.org/series/SP500)

We found our data for the building permits came from the Building Permits Survey from the US Census data. This dataset provides data on the number of new housing units authorized by building permits. 
The construction material index was sourced from the Federal Reserve Bank of St. Louis the FRED.

[BPS](https://www.census.gov/construction/bps/)
[Construction Materials Index](https://fred.stlouisfed.org/series/WPUSI012011)


# US National Sale Volume Changes

```{r, echo=FALSE}
linegraph_data %>%
ggplot(aes(x=date, y=sales_count, color = region)) +
  geom_line() +
  geom_point() +
  ggtitle("Home Sales Count by US Region over Time") +
  labs(x = "Date", y = "Sales Count", color = "Region of the US")
```

We tracked the patterns in national sale volume over time, from July 2019 to March 2021. From this visualization, we can see that national sales hit a sharp downturn and reached its lowest point in Feb-March 2020, then began to spike back up in April and into the summer. The first trend can be explained by the nationwide lockdown, as all businesses (e.g. real estate agencies) were forced to shut down or operate remotely, and fear about the pandemic and its transmission were at an all-time high. The second trend can be explained by the “suburban shift” and exodus to low-tax states, as the remote economy no longer required proximity to urban centers and encouraged workers to move to larger homes. These trends can be seen across different regions of the US - each of the four lines follow a similar pattern. Such similarity indicates that these trends are nationwide, and thus likely caused by nationwide occurrences (like a global pandemic). However, we see another downturn in Jan 2021, as surging home prices (due to low supply and high demand) made home purchases less affordable for the average consumer. 

# US Average Home Price Changes

```{r, echo=FALSE}
#Create the box plot
BoxPlot <- BoxPlotData %>%
  ggplot(aes(x=region, y=price, fill=region)) +
    geom_boxplot() +
  #Create a box plot for each of the three dates
    facet_wrap(~date, labeller = label_both) +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11),
    ) +
    ggtitle("Housing Prices by Region, Year by Year") +
    xlab("Region") +
    theme_ipsum()

#turn the text on x-axis 50 degrees
BoxPlot <- BoxPlot + theme(axis.text.x = element_text(angle = 50, hjust = 1))


BoxPlot
```

# Home Price by State, Interactive Spatial Map

```{r, echo=FALSE}
#Create the table that holds the data to create my map
  MapTable <- FullTableLong %>%
    filter(date == "X2021.03.31" | date == "X2020.03.31")

#create a variable that calculates the percent change from 2020 to 2021
  MapTableNext <- MapTable %>%
    group_by(RegionName) %>%
      mutate(PercentChange = round(((price - lag(price))/lag(price))*100, 3)) %>%
      filter(date == "X2021.03.31")

  #Create the variable that holds US state map data
usa_states <- map_data(map = "state"
                       , region = ".")
#Revise the table to only include the data points wanted for the map
  MapTable <- MapTable %>%
    filter(date == "X2021.03.31")

#Mutate the data names in the MapTable so they line up with the names in the
  #leaflet
MapTable <- MapTable %>%
  mutate("RegionName" = tolower(RegionName))

#creating new dataset that joins the graphs together
MapTable2 <- MapTable %>%
  inner_join(usa_states, by = c("RegionName" = "region"))

#Mutate the data names in the MapTableNexr so they line up with the names in the
#leaflet
MapTableNext <- MapTableNext %>%
  mutate("RegionName" = tolower(RegionName))
options(scipen=10000)

#Create the leaflet and assign the names
usa_states_leaflet <- map("state", fill = TRUE, plot = FALSE)
class(usa_states_leaflet)
usa_states_leaflet$names

#Takes all state names and puts them in a dataframe
usa_states_leaflet2 <- as.data.frame(usa_states_leaflet$names) %>%
#Renames variable for clarity
  rename(RegionName = 'usa_states_leaflet$names') %>%
#Isolates the state name and discards information regarding region within state
  separate(col = RegionName, into = c("RegionName2", "extra"), sep = ":",
           remove = FALSE)

#Combines "blank map" with data on housing prices
FullMap <- usa_states_leaflet2 %>%
  left_join(MapTableNext, by= c("RegionName2" = "RegionName"))

#Sets up the color palate for the map
mypal <- colorNumeric(
  palette = "YlGnBu",
  domain = MapTable$price
)

#creates the leaflet map of housing prices in the US
PriceMap <- leaflet(data = usa_states_leaflet) %>%
  addTiles() %>%
  setView(-90,42.4, zoom = 3.5) %>%
  # base state's fill color on housing price
  addPolygons(fillColor = ~mypal(FullMap$price)
              , color = "#FFFF33"  #hex format of color
              , fillOpacity = 0.9
              , stroke = FALSE
              #determines what will pop up when a state is clicked
              , popup = paste0("State: ", FullMap$RegionName2, "<br>"
                             , "Average Price as of March, 2021: $", FullMap$price, "<br>"
                             ,"Year over year price change: ", FullMap$PercentChange, "%<br>"
                              )) %>%
  # add legend explaining fill color
   addLegend(pal = mypal,
            values = FullMap$price,
            position = "bottomright",
            title = "Home Prices by State as of March 31, 2021")

PriceMap
```

# Comparison of Housing Price Index, Consumer Price Inde, and S&P 500

```{r, echo=FALSE}
Indexplot <- CPIHPI %>%
  ggplot(aes(x = Date, y = Index_Values, group = Index_Type, color = Index_Type)) +
  geom_line() +
  geom_point() +
  ylab("Index Values") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(color = "Index Type") +
  ggtitle("Consumer Price Index vs. Housing Price Index")

Indexplot
```

The HPI has outperformed the CPI over the course of 20 years and certainly stands as an asset class that appreciates faster than inflation. Although we do see a large rise and fall between 2005-2012, we can see that the HPI has recovered even higher than the past peak. The housing market thus seems like a more volatile compared to the living expenses of the average US citizen. This indeed allows us to realize that investing in real estate is more likely than not to grow and is a better alternative to holding cash. However, we also look to see how real estate has performed compared to the stock market in the form of the S&P 500 index. 

```{r, echo=FALSE}
SPplot <- SPHomePrice %>%
  ggplot(aes(x=Date, y=pct_chg, group = Index_Type, color = Index_Type)) +
  geom_line() +
  ylab("Percentage Change of Index") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(color = "Index Type") +
  ggtitle("S&P 500 vs. Case Shiller Housing Price Index")

SPplot
```

In comparison with the S&P 500, we can see that the HPI is much more stable compared to the volatility of the stock market. While the HPI averages around 3%, we can see that the S&P 500 grows at a 7% average. In this case, we can see that the top 500 stocks will perform better than the average house in the US. However, there are some caveats that need attention. First is that the S&P 500 tracks the top 500 stocks while the HPI tracks all home values. Thus, it is not the best comparison when comparing returns due to the fact the HPI considers personal residencies that do not try to achieve the returns of investment properties, while all S&P 500 companies are trying to achieve a high return. 

The difference in volatility may be explained in the liquidity of the market. Stock are able to be bought and sold in a relatively quick manner while real estate may take weeks, months, or even years to buy or sell properties. Thus, market changes are much more pronounced in the stock market while changes may be slow to show in the real estate market. 

# Interest and Home Price Relationship

```{r, echo=FALSE}
ggplot(data=scatter_data, mapping = aes(x = interest_rate, y = median_price_thousand)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  ggtitle("Interest Rate vs Home Price in the United States") +
  scale_y_continuous(labels = function(x) paste0("$", x)) +
  labs(x = "Mortgage Interest Rates (%)", y = "Median Sale Price (in Thousands)")
```

The downward-sloping line of best fit demonstrates the negative correlation between interest rates and home price. Looking at monthly data since 2008, as interest rates rose, median home price generally fell. We can infer that this is due to the attractiveness of spending and investing during periods of low interest rates - loans for large investments like real estate will accrue less interest and therefore prove less costly over time during periods of low interest rates. Thus, demand for real estate increases, pushing up prices. This relationship can help explain the surge in the real estate market during the pandemic - as interest rates fell, investments in real estate became more attractive, especially during a time when remote work and stay-at-home orders encouraged moving to larger spaces with lower taxes.

# Construction Material Costs and Building Permits Changes

```{r, echo=FALSE}
BPvCMplot <- BPvCM %>%
  ggplot(aes(x=Date, y=pct_chg, group = Data, fill = Data)) +
  geom_area() +
  ylab("Percentage Change") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(fill = "Data") +
  ggtitle("Difference between Building Permits Approved and Construction Material Price")

BPvCMplot
```

Interestingly, we actually see that building permits increase when construction material has increased. This is a perplexing finding as we expect building permits to increase as construction material costs decrease. We expected this outcome because homebuilders can expect a higher return when the costs of building decrease, which incentivizes more home builders to build. Yet this was not the case. This may be because although construction material prices increased, sales price also increased accordingly, which invited more new builds. In addition, because the building permits are approved by the government, there may be a delay from when a home builder applies to build to when a home builder receives a permit. 

# Conclusion and Limitations

## Answering Questions of Interest

**1. How did the COVID-19 pandemic impact the real estate industry?**

We observed that sales prices saw a slight dip, but rebounded and actually recorded record growth. Overall sales count also increased and the real estate market has seen strong growth throughout the pandemic. This was contrary to many people's belief that the market was to crash like 2008. However, we see that foreclosure protection as well as eviction protection, allows the market to stay robust. 

**2. How well has the real estate market performed compared to other markets?**

We found that, despite the pandemic, real estate has performed marketedly well in comparison to other assets. In comparison to the CPI, we were able to observe that real estate historically beat out inflation. However, in comparison to the S&P 500, it lacked the returns. However, it was much less volatile than stocks during distressed times as a stable investment. However, this may be due to the forbearance and foreclosure protection, so it will be interesting to see how the market develops once these protections have dissappeared.

Yet this comparison does not factor in the benefits of owning real estate and the adjusted returns incorporating such benefits. The HPI shows the average appreciation of homes, but does not include the depreciation, tax benefits, cash flow, principal paydown, leverage, and forced appreciation that homeowners enjoy.

This also does not include the 1031 exchange, which allows homeowners defer their taxes to later years and sell their property to buy a greater or equal value home. This in turn allows the investor to invest pre-tax cash to achieve higher returns with a bigger bankroll.

Thus, while some investors may consider real estate an inferior investment to stocks based on these indexes, but fail to realize the unrecorded benefits of real estate.

**3. What explains this real estate market behavior?**

Our findings on the negative correlation between interest rate and sale price indicate that lower interest rates lead to higher demand for real estate and therefore higher sale prices. Thus, the Fed's response in early 2020 to support the US economy during the lockdown by slashing down to near-zero interest rates could explain the surge in sales count and sales price (Cheng et al. 2021).

We actually found evidence against the theory that construction material price increases decrease the amount of homebuilders trying to build a home. This was a perplexing finding, but we were able to realize that there may be delays of when builders apply and get approved to build homes. 

## Limitations

One limitation to our study is that the real estate market does not have perfect information and includes a large off-market section. We are unable to track these transactions and thus does not provide the full picture of all real estate transactions. However, we are able to track those listed on MLS and sites such as Zillow and Redfin, which allow us to graspe a general sense of the real estate market. 

Our construction data and findings could also be limited by government bureaucracy. Government and market move at different paces, the government might approve permits at different dates than the application. Thus, building permits approved may not yet accurately reflect the larger claims we believe to be true for supply of homes.

## Possible Next Steps

Possible next steps would to understand the debt markets of real estate and see the impact of covid on mortgages. For instance, we can track missed and late payment percentages and see how Covid impacted those numbers. In addition, we could see how the total mortgage debt of the US was impacted by covid and if the mortgage growth rate was irregular during this period. 

Another angle that would be interesting to look at are different asset classes. We focused on the SFH market, but there are various other asset classes such as multifamily, industrial, retail, hotels, land, and office. However, publicly available data is limited so a subscription to Loopnet or a commercial Multiple Listing Service (MLS) would be extremely helpful in this instance. 

Although this was an overall analysis of the US real estate market, geography plays a massive part. Thus, it may be more interesting to look into specific states, counties, and neighborhoods to understand how the pandemic has affected local markets. 

# Citations

Odeta Kushi. "Homeownership Remains Strongly Linked to Wealth-Building." Nov 5, 2020.  https://blog.firstam.com/economics/homeownership-remains-strongly-linked-to-wealth-building/ 

Cheng et al. "What’s the Fed doing in response to the COVID-19 crisis? What more could it do?" March 30, 2021 https://www.brookings.edu/research/fed-response-to-covid19/
