---
title: "Josh_DataWrangling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(robotstxt)
library(knitr)
library(janitor)
library(rvest)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(leaflet)
library(leaflet.extras)
library(maps)
```

## HPI vs CPi and SP&500 vs Case Shiller HPI

```{r}
#Import CPI data
CPI <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/USACPIALLMINMEI.xls", range = "A11:B265") %>%
    #Renaming columns to make sense
  rename(Date = "observation_date", ConsumerPriceIndex = "USACPIALLMINMEI") %>%
    #formatting the class of this column for visualization
  mutate(Date = as.Date(as.character(Date), format = "%Y-%m-%d"))
  
#Import HPI
HPI <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/CSUSHPINSA.xls", range = "A11:B265") %>%
    #Renaming columns to make sense
  rename(Date = "observation_date", HousingPriceIndex = "CSUSHPINSA") %>%
    #formatting the class of this column for visualization
  mutate(Date = as.Date(as.character(Date), format = "%Y-%m-%d"))

#Join two tables together
CPIHPI <- HPI %>%
  left_join(CPI, by = "Date") %>%
  #pivot to long dataset for visualization
  pivot_longer(!Date, values_to = "Index_Values", names_to = "Index_Type")

#Import S&P 500 data with Case Shiller Home Price
SPHomePrice <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/sp-500-vs-home-prices.xls.xlsx") %>%
  #Rename columns to legible names
  rename(DateTime = "S&P 500 vs. Home Prices", "SP_500" = "...2", "Case_Shiller_Home_Price_Index" = "...3") %>%
  #Filter for 2 years worth of data from 2019-2021
  filter(DateTime > "2019-02-01") %>%
  #filtering out one row in which does not contain observation
  filter(DateTime != "DateTime") %>%
  #formattting columns to make sure they work in computations and visualizations
  mutate(
    Date = as.Date(DateTime, format = "%Y-%m-%d"),
    MY = format(Date, format = "%Y-%m"),
    SP_500 = as.double(SP_500),
    Case_Shiller_Home_Price_Index = as.double(Case_Shiller_Home_Price_Index)
    ) %>%
  #removing unnecessary column
  select(-DateTime) %>%
  #Pivot to long dataset for visualization
  pivot_longer(!Date & !MY, names_to = "Index_Type", values_to = "Index_Value") %>%
  #filtering out NA values
  filter(Index_Value != "NA") %>%
  #Group by month and year to find datapoint for last day of each month
  group_by(MY) %>%
  filter(Date == max(Date)) %>%
  #ungroup to allow computations for percentage change
  ungroup() %>%
  #arrange by index type and date to order data correctly
  arrange(Index_Type, Date) %>%
  #Calculate percentage change
  mutate(
    pct_chg = as.double(100 * (Index_Value - lag(Index_Value))/lag(Index_Value))
         ) %>%
  #Change initial percentage change to 0 for visualization
  within(pct_chg[Date == '2019-02-28'] <- as.numeric('0'))

```

## Building Permit vs Construction Material Index Percentage Change

```{r}
#Import exel sheet with observations only
BP202002 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202002.xls", range = "A8:G74") %>%
  #rename column for legibility
  rename(Region = "...1") %>%
  #filter out NA observations and non-state observations
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  #pivot to long data for visualization
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  #change class of column for visualization
  mutate(Date = as.Date("02-01-2020", "%m-%d-%Y"))

#Continuted for one whole year
BP202003 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202003.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("03-01-2020", "%m-%d-%Y"))

BP202004 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202004.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("04-01-2020", "%m-%d-%Y"))

BP202005 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202005.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("05-01-2020", "%m-%d-%Y"))

BP202006 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202006.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("06-01-2020", "%m-%d-%Y"))

BP202007 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202007.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("07-01-2020", "%m-%d-%Y"))

BP202008 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202008.xls", range = "A8:G74") %>%
 rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("08-01-2020", "%m-%d-%Y"))

BP202009 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202009.xls", range = "A8:G74") %>%
 rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("09-01-2020", "%m-%d-%Y"))

BP202010 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202010.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("10-01-2020", "%m-%d-%Y"))

BP202011 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202011.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("11-01-2020", "%m-%d-%Y"))

BP202012 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202012.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("12-01-2020", "%m-%d-%Y"))

BP202101 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202101.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("01-01-2021", "%m-%d-%Y"))

BP202102 <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/statemonthly_202102.xls", range = "A8:G74") %>%
  rename(Region = "...1") %>%
  filter(Total != "NA") %>%
  filter(!Region == "Northeast Region") %>%
  filter(!Region == "New England Division") %>%   
  filter(!Region == "Middle Atlantic Division") %>%   
  filter(!Region == "Midwest Region") %>%   
  filter(!Region == "East North Central Division") %>%   
  filter(!Region == "West North Central Division") %>%   
  filter(!Region == "South Region") %>%   
  filter(!Region == "South Atlantic Division") %>%   
  filter(!Region == "East South Central Division") %>%   
  filter(!Region == "West South Central Division") %>%   
  filter(!Region == "West Region") %>%   
  filter(!Region == "Mountain Division") %>%   
  filter(!Region == "Pacific Division") %>%
  pivot_longer(cols = !Region, names_to = "Housing_Type", values_to = "Count") %>%
  mutate(Date = as.Date("02-01-2021", "%m-%d-%Y"))

#Building Permits binding together into one dataset
BuildingPermits <- bind_rows(BP202002, BP202003, BP202004, BP202005, BP202006, BP202007, BP202008, BP202009, BP202010, BP202011, BP202012, BP202101, BP202102) %>%
  #Filter for total US numbers
  filter(Housing_Type == "Total") %>%
  #Column no longer needed
  select(-Housing_Type) %>%
  #filter for US total numbers
  filter(Region == "United States") %>%
  #Calculate percentage change and add Data column to merge with ConstructionMaterial
  mutate(
    pct_chg = as.numeric(100 * (Count - lag(Count))/lag(Count)),
    Data = "Building Permits"
         ) %>%
  #set initial percentage change to 0
  within(pct_chg[Date == '2020-02-01'] <- as.numeric('0')) %>%
  #remove unneccessary columns
  select(-Count, -Region)
  

#Import Construction Material Index Data
ConstructionMaterial <- read_excel("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/Joshua/PCU44414441.xls", range = "A23:B37") %>%
  #Rename columns for legibility
  rename(
    Date = "43831",
    Material = "153.5"
    ) %>%
  #calculate percentage change and add column to join with BuildingPermits
  mutate(
    pct_chg = as.numeric(100 * (Material - lag(Material))/lag(Material)),
    Data = "Construction Material"
  ) %>%
  #set initial percentage change to 0
  within(pct_chg[Material == "152.2"] <- as.numeric('0')) %>%
  #Remove unncessary columns
  select(-Material)

#Bind these two together for visualization
BPvCM <- ConstructionMaterial %>%
  bind_rows(BuildingPermits)

#Create csv files for easy importation into visualizations
write_csv(BPvCM, file="Datasets/BPvCM.csv")
write_csv(CPIHPI, file="Datasets/CPIHPI.csv")
write_csv(SPHomePrice, file="Datasets/SPHomePrice.csv")
```

# Visualizations
```{r}
SPplot <- SPHomePrice %>%
  ggplot(aes(x=Date, y=pct_chg, group = Index_Type, color = Index_Type)) +
  geom_line() +
  ylab("Percentage Change of Index") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(color = "Index Type") +
  ggtitle("S&P 500 vs. Case Shiller Housing Price Index")
            
SPplot

Indexplot <- CPIHPI %>%
  ggplot(aes(x = Date, y = Index_Values, group = Index_Type, color = Index_Type)) +
  geom_line() +
  geom_point() +
  ylab("Index Values") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(color = "Index Type") +
  ggtitle("Consumer Price Index vs. Housing Price Index")

Indexplot

BPvCMplot <- BPvCM %>%
  ggplot(aes(x=Date, y=pct_chg, group = Data, fill = Data)) +
  geom_area() +
  ylab("Percentage Change") +
  theme(legend.position="bottom", legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid")) +
  labs(fill = "Data") +
  ggtitle("Difference between Building Permits Approved and Construction Material Price")

BPvCMplot
```

#Ethan 
```{r}
#download state by state housing prices
valueTableState<- read.csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Zillow_homevalue.csv")

#download national and city housing prices
valueTableUS <- read.csv("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Zillow_USMetro.csv")

#select certain dates for price data, and only include the US data, none of the cities
valueTableUS2 <- valueTableUS %>%
  select(3,5,282:308) %>%
  filter(RegionName == "United States")

#select certain dates for price data after 2019
valueTable2 <- valueTableState%>%
  select(3,5,282:308)

#Combine the overall US data with state by state data
FullTable <- valueTable2 %>%
  full_join(valueTableUS2)

#Pivot the data frame to a long data style
  FullTableLong <- FullTable %>%
  pivot_longer(c(-"RegionName",-"StateName"), names_to = "date" , values_to = "price")
  
  #Create a table that only has US data
  USTable <- FullTableLong %>%
    filter (RegionName == "United States")
  
write_csv(FullTableLong, file="Datasets/FullTableLong.csv")
```


```{r}
#Categorize each state within the region they are in
northeast <- c("NY", "NJ", "PA", "CT", "ME", "MA", "NH", "RI", "VT")
midwest <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MO", "MN", "NE", "ND", "SD")
south <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "DC",
          "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
west <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")

#filter the data to include the date that I want to include in the box plot
BoxPlotTable <- FullTableLong %>%
  filter(date == "X2019.03.31" | date == "X2020.03.31" | date == "X2021.03.31")

#Add a variable that includes which region each state is in
BoxPlotData <- BoxPlotTable %>%
  mutate(
    region = case_when(StateName %in% northeast ~ "NORTHEAST",
                            StateName %in% midwest ~ "MIDWEST",
                            StateName %in% south ~ "SOUTH",
                            StateName %in% west ~ "WEST",
                            RegionName == "United States" ~ "US",
                            TRUE ~ "")
    ) %>%
  rename(city = RegionName,
         state = StateName) %>%
  mutate_at("date", str_replace, "X", "")

#change the axis from scientific notation to normal numbers
options(scipen=10000)

#Create the box plot
BoxPlot <- BoxPlotData %>%
  ggplot(aes(x=region, y=price, fill=region)) +
    geom_boxplot() +
    facet_wrap(~date) +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Housing Prices by Region, Year by Year") +
    xlab("Region")  +
    ylab("Price")

#turn the text on x-axis 50 degrees
BoxPlot <- BoxPlot + theme(axis.text.x = element_text(angle = 50, hjust = 1))

  BoxPlot

write_csv(BoxPlotData, file="Datasets/BoxPlotData.csv")
```


```{r}
#Create the table that holds the data to create my map
  MapTable <- FullTableLong %>%
    filter(date == "X2021.03.31" | date == "X2020.03.31")

#create a variable that calculates the percent change from 2020 to 2021
  MapTableNext <- MapTable %>%
    group_by(RegionName) %>%
      mutate(PercentChange = round(((price - lag(price))/lag(price))*100, 3)) %>%
      filter(date == "X2021.03.31")

  #Create the variable that holds US state map data
usa_states <- map_data(map = "state"
                       , region = ".")
#Revise the table to only include the data points wanted for the map
  MapTable <- MapTable %>%
    filter(date == "X2021.03.31")

#Mutate the data names in the MapTable so they line up with the names in the
  #leaflet
MapTable <- MapTable %>%
  mutate("RegionName" = tolower(RegionName))

#creating new dataset that joins the graphs together
MapTable2 <- MapTable %>%
  inner_join(usa_states, by = c("RegionName" = "region"))

#Mutate the data names in the MapTableNexr so they line up with the names in the
#leaflet
MapTableNext <- MapTableNext %>%
  mutate("RegionName" = tolower(RegionName))
options(scipen=10000)

#Create the leaflet and assign the names
usa_states_leaflet <- map("state", fill = TRUE, plot = FALSE)
class(usa_states_leaflet)
usa_states_leaflet$names

#Takes all state names and puts them in a dataframe
usa_states_leaflet2 <- as.data.frame(usa_states_leaflet$names) %>%
#Renames variable for clarity
  rename(RegionName = 'usa_states_leaflet$names') %>%
#Isolates the state name and discards information regarding region within state
  separate(col = RegionName, into = c("RegionName2", "extra"), sep = ":",
           remove = FALSE)

#Combines "blank map" with data on housing prices
FullMap <- usa_states_leaflet2 %>%
  left_join(MapTableNext, by= c("RegionName2" = "RegionName"))

#Sets up the color palate for the map
mypal <- colorNumeric(
  palette = "YlGnBu",
  domain = MapTable$price
)

#creates the leaflet map of housing prices in the US
PriceMap <- leaflet(data = usa_states_leaflet) %>%
  addTiles() %>%
  setView(-90,42.4, zoom = 3.5) %>%
  # base state's fill color on housing price
  addPolygons(fillColor = ~mypal(FullMap$price)
              , color = "#FFFF33"  #hex format of color
              , fillOpacity = 0.9
              , stroke = FALSE
              #determines what will pop up when a state is clicked
              , popup = paste0("State: ", FullMap$RegionName2, "<br>"
                             , "Average Price as of March, 2021: $", FullMap$price, "<br>"
                             ,"Year over year price change: ", FullMap$PercentChange, "%<br>"
                              )) %>%
  # add legend explaining fill color
   addLegend(pal = mypal,
            values = FullMap$price,
            position = "bottomright",
            title = "Home Prices by State as of March 31, 2021")

PriceMap

write_csv(usa_states_leaflet, file="Datasets/usa_states_leaflet.csv")
write_csv(FullMap, file="Datasets/FullMap.csv")
```



## National Sales Volume Difference over Time (Box Plot)

```{r}
#get Salecount data. Want the points for different cities, no average. 
#remove unnecessary columns
SalesCountData <- read_csv(paste0("/home/class21/dchoe21/git/Blog-Funemployed-Statisticians/Datasets/sales_count.csv"))

SalesCountData <- SalesCountData %>%
  filter(SizeRank != 0) %>%
  select(-"RegionID", -"SizeRank", -"RegionType")

#assign states to regions
northeast <- c("NY", "NJ", "PA", "CT", "ME", "MA", "NH", "RI", "VT")
midwest <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MO", "MN", "NE", "ND", "SD")
south <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "DC",
          "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
west <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")

SalesCountStateData <- SalesCountData %>%
  mutate(region = case_when(StateName %in% northeast ~ "NORTHEAST",
                            StateName %in% midwest ~ "MIDWEST",
                            StateName %in% south ~ "SOUTH",
                            StateName %in% west ~ "WEST",
                            TRUE ~ "")) %>%
  rename(city = RegionName,
         state = StateName)
        
#pivot for different date values
SalesCountDataLong <- SalesCountStateData %>%
  pivot_longer(c(-"city", -"state", -"region"), names_to = "date", values_to = "sales_count")%>%
  mutate(date =  as.Date(date, format= "%Y-%m-%d")) 


linegraph_data <- SalesCountDataLong %>%
  filter(date >= as.POSIXct("2019-06-30")) %>%
  group_by(date, region) %>%
  summarize(sales_count = mean(sales_count, na.rm = TRUE))

write_csv(linegraph_data, file="Datasets/linegraph_data.csv")
```

## Interest vs Home Price (Scatter Plot)

```{r}
#Get FRED Data
FRED_Data <- read_csv(paste0("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/FRED_interest.csv")) %>%
  rename(date = DATE, interest_rate = MORTGAGE15US) %>%
  mutate(date =  as.Date(date, format= "%Y-%m-%d"), 
         year = year(date),
         month = month(date)) %>%
  filter(date > "2008-04-30") %>%
  group_by(month, year) %>%
  summarize(interest_rate = mean(interest_rate))

#Zillow Data on median home price monthly. Want US average
MonthlySalePriceData <- read_csv(paste0("/home/class21/jokim21/Git/Blog-Funemployed-Statisticians/Datasets/median_monthly.csv"))
USaverageprice <- MonthlySalePriceData %>%
  filter(SizeRank == 0) %>%
  select(-"RegionID", -"SizeRank", -"RegionName", -"RegionType", -"StateName") %>%
  pivot_longer(col = everything(), names_to = "date", values_to = "median_price") %>%
  mutate(date = as.Date(date, format= "%Y-%m-%d"),
         year = year(date),
         month = month(date),
         median_price_thousand = median_price / 1000)


#combine the two datasets
scatter_data <- USaverageprice %>%
  inner_join(FRED_Data, by = c("year", "month")) %>%
  select(-"date")

write_csv(scatter_data, file="Datasets/scatter_data.csv")
```

# VISUALIZATIONS

## National Sales Volume Difference over Time (Line Plot)

Problem is that we have several points for the same region on the same date. Use groupby and summarize to average out the values so there is only one point per x-value per region, and also figure out what's going on with the date limit to after June 2019. 
```{r}
linegraph_data%>%
ggplot(aes(x=date, y=sales_count, color = region)) +
  geom_line() +
  geom_point() +
  ggtitle("Home Sales Count by US Region over Time") +
  labs(x = "Date", y = "Sales Count", color = "Region of the US")
```

##Interest vs Home Price (Scatter Plot)

```{r}
ggplot(data=scatter_data, mapping = aes(x = interest_rate, y = median_price_thousand)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  ggtitle("Interest Rate vs Home Price in the United States") +
  scale_y_continuous(labels = function(x) paste0("$", x)) +
  labs(x = "Mortgage Interest Rates (%)", y = "Median Sale Price (in Thousands)")
```
